Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law

### 2023-12-04 manuscript/manuscript.qmd
Minor revisions + update legal section + add coding rules
Resolved some of the TO DOs

---
title: "The Operational Space of International Civil Society Under Authoritarian Regulations: The Case of China"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-12-03"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

In the past two decades, nearly 100 countries have proposed or enacted legal measures designed to restrict nongovernmental organizations (NGOs) and civil society more broadly [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016].[^ngo-clarification] According to the CIVICUS Civil Society Monitor, as of 2023 only `r civicus_prop$true$prop_nice` of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @CIVICUSMonitor]. This global trend of closing or shrinking civic space has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of civil society restrictions with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China---or "ONGO Law"---which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either establish formal in-country representative offices (ROs) or file short-term temporary activities (TAs) to lawfully operate in mainland China. Without recognized legal status, INGOs are prohibited from working in the country, and even with legal recognition, INGO programming is limited to government-approved activities.

@@

By focusing on INGO operational space in China, we provide a unique view into the *de facto* regulatory environment for foreign civil society in an authoritarian state. Many previous studies use broad indexes of civil society openness, such as the core civil society index from the Varieties of Democracy project [@LianMurdie:2023; @ChaudhryHeiss:2022]. These measures capture the general regulatory environment for all civil society---including domestic NGOs, INGOs, and informal organizations---but do not reflect the specific implementation of laws for INGOs. In this paper, though, we use a mixed methods approach to explore the regulatory environment for INGOs in particular, providing much a richer description of authoritarian regulation of international civil society.

Below, we provide a brief overview of China's 2017 ONGO Law and the different regulatory mechanisms---and practical limitations---it provides for INGOs to obtain legal access to the country. We then summarize existing research on authoritarian regulation of INGOs and the differential application of these laws based on regime preferences and organizational characteristics, and propose a set of empirical expectations based on this literature. We test these expectations with a mixed methods research design, combining Bayeisan analysis of administrative data from `r nrow(ongo)` formally registered INGOs with a comparative case study of Greenpeace and The Nature Conservancy, two INGOs working on similar issue areas which took different approaches to operating in China after 2017. We conclude with a broader discussion of authoritarian civil society regulations and propose future methods for exploring the practical operating environment for INGOs working in such regimes.

# The legal framework established by the 2017 ONGO law

International NGOs began to play an important role in China in the early 1990s as part of the country's Reform and Opening-Up policy, and organizations such as Save the Children and the Ford Foundation established local offices and branches. Since then, INGOs and global civil society have played pivotal roles in providing funding, organizational and program management skills, as well as catalytic support for the emergence of China's domestic nonprofit sector [@Yin:2009]. Prior to the implementation of the 2017 ONGO Law, regulation of INGOs in China was ambiguous [@Shieh:2018; @Spires:2022]. The only legal document relevant to INGOs was a section of the Regulation on the Management of Foundations by the State Council that specifies the registration of representative offices ("ROs") of foreign foundations with the Ministry of Civil Affairs, the regulator of Chinese domestic NGOs. However, the actual application of the regulation was very limited: until the ONGO Law took effect in 2017, there were fewer than 30 foreign foundation ROs registered at the Ministry of Civil Affairs, while some Chinese officials estimated in 2016 that there were roughly over 7,000 INGOs operating in China [@YeHuang:2018; @Spires:2020].

Many INGOs gained legal status within mainland China by registering as foreign-funded companies [@Sidel:2016], though there were no clear requirements to have a formal legal presence in China [@Han:2011; @HolbigLang:2022; @Yin:2009]. However, the legal form of foreign-funded company is not quite compatible with the non-profit-distributing nature of INGOs. Under the Chinese legal system, enterprises were intended to be for-profit entities while nonprofit organizations were required to choose other standalone legal forms, such as foundations, associations, and social service organizations [@Ye:2021]. Thus, INGOs obtaining legal status had to navigate a legal framework designed for a for-profit institutional logic, which often conflicted with nonprofit characteristics like their lack of private investment returns.

The legal framework governing INGOs was overhauled by the 2017 Chinese ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) non-governmental organizations that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief. The ONGO Law provides two pathways for INGOs to lawfully conduct nonprofit activities in China (see @fig-ro-ta-pathways): (1) submitting an application for formal permanent registration as ROs, or (2) filing documentation of temporary activities (TAs) with the Ministry of Public Security and provincial-level Public Security Bureaus. Formal registration as an RO allows INGOs to maintain an ongoing presence in the country, while TAs are intended for INGOs that need to conduct one-off programs that last no longer than one year. Using one of these two channels to obtain legal status became mandatory in 2017 and any INGOs that could not successfully obtain proper legal authorization were deemed illegal.

@@

ro_nodes <- tibble(
  step = c(
    "Find a professional supervisory unit (PSU)",
    "Obtain PSU approval",
    "Open organization bank accounts",
    "Submit annual work plans to PSU and Public Security Bureau",
    "Keep programming within approved scope",
    "Submit annual reports to PSU and Public Security Bureau"
  ),
  type = "Registered organization (RO)"
)
ro_edges <- tibble(from = 1:5, to = 2:6)
ro_graph <- tbl_graph(nodes = ro_nodes, edges = ro_edges)

ta_nodes <- tibble(
  step = c(
    "Find an eligible Chinese partner to file paperwork",
    "File activity information with Public Security Bureau",
    "Keep programming within approved scope",
    "Finish work within a year",
    "Submit final project report to Public Security Bureau"
  ),
  type = "Temporary activities (TA)"
)
ta_edges <- tibble(from = 1:4, to = 2:5)
ta_graph <- tbl_graph(nodes = ta_nodes, edges = ta_edges)

@@

Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. Organizations vary in their ability to attract partner PSUs, and INGOs that are unable to find suitable PSUs can end up not being able to registered or registered with great limitation in terms of work areas and geographical reach. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province.

@@

The authoritarian benefits of civil society regulations go beyond simply allowing or disallowing specific INGOs. The flexible *de facto* implementation of formal civil society laws allows authoritarians to "use nonprofit regulation as a tool of political control to shore up their continued rule" [@Spires:2020, 573], providing them with an additional policy tool for their stability-seeking calculus. These laws allow governments to capture and reshape INGO programming in their favor. @LianMurdie:2023 illustrate how increased civil society restrictions have counterintuitively led to an increase in cooperative, regime-friendly INGO programming, with INGOs praising government policies, engaging in joint programming with the government, and changing longstanding programs to align with government desires. China's ONGO law illustrates this dynamic well. The ONGO law requires that INGOs partner with provincial government agencies and find collaborating PSUs, which provides the government with more influence over INGO decision making. For example, Oxfam gained RO status in 2017, but only after negotiations with its proposed PSU forced it to stop collaborating with other organizations and local agencies [@Li:2020]. In 2018, @BatkeHang:2018 posited that the ONGO law was showing signs of reorienting the missions of legally recognized INGOs and the overall scope of international civil society to be more explicitly aligned with government preferences. Holbig & Lang confirmed this prediction, finding that the issue areas addressed by legally recognized INGOs had "shift\[ed\] towards fields of activity high up the government's domestic policy agenda" [-@HolbigLang:2022, p. 587]. Civil society laws thus filter out organizations opposed to the host regime's policy preferences and grant increased government control over safer and more beneficial INGOs.

@@

To test the relationships between government preferences, INGO issue area, and INGO operational space, we apply this framework to China's ONGO law. Previous research has looked at the practical implementation of civil society laws on a global scale [@LianMurdie:2023; @Chaudhry:2022], which naturally blurs the specific policy mechanisms regimes use for controlling and leveraging global civil society. By analyzing the population of formally registered INGOs between 2017–2022, we can more precisely explore how a single formal anti-INGO law has shaped the operational space for foreign organizations working in the country. Specifically, we can observe exactly how the Chinese government has used the RO registration process to strategically limit the potential reach of INGOs and reshape their issue areas. We use the number of provinces that INGOs are authorized to work in as a proxy for INGO operational space based on the assumption that INGOs need to or hope to be able to carry out their programs in as broad a geographical scope as possible [@Li:2020]. For example, in emergency circumstances like the COVID-19 pandemic, it would be extremely difficult for INGOs to plan beforehand which areas might need their service. Having the potential to shift programs to other provinces as needed thus provides a form of operational space.

@@

We examine these theoretical expectations using a mixed methods approach. First, we explore a rich dataset of all ROs registered in China between 2017–2021 using a combination of descriptive statistics and Bayesian models to understand how the international nonprofit sector in China has been reshaped since the implementation of the ONGO law. We then complement this statistical analysis with a comparative case study of two INGOs---Greenpeace and The Nature Conservancy---to explore in detail how the ONGO has specifically influenced organizational operational space.

@@

Our data comes from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office, which maintains a website with complete documentation of all the ROs and temporary activities registered or filled across China. We include all INGO registrations between January 1, 2017, when the law took effect, to December 31, 2021, with information for all `r nrow(ongo)` registered organizations. The filing documents for ROs disclose each INGO's name, address, registration date, regulating Public Security Bureau, professional supervisory unit (PSU),  issue area or purposes of the RO, and the specific provinces they can work in. We merge this INGO data with translated English organization names data from ChinaFile. Our data and reproducible code is available at `ANONYMIZED_DOI_URL`.

@@

Our dependent variable is the count of provinces that INGOs are authorized to work in. We derive this value from the official registration record, which lists all the provincial regions each INGO RO can conduct activities in. This value ranges from 1–32 (the total number of Chinese provinces).

Our key independent variables correspond to each of our three theoretical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, we cross-checked our manual coding of issue areas---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix (TODO). We categorize each organization into nine different issue areas (see @tbl-issues).

@@

Our second independent variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally or subprovincially. There are cases where the aim of the INGO is attached to a locality, either having a county or city name in their RO name or purposes, or the INGO is actually set up by people in China who registered the RO to communicate between the "hometown" and the INGO set up abroad.

Our final independent variable measures how much time elapsed between January 1, 2017 (when the law took effect) and the organization's official registration date. There is great variance in the timing of registering an RO among INGOs, since there has been a *de facto* grace period for INGOs to seek legal presence after the law took effect [@Shieh:2018]. This is largely because both INGOs and the regulators needed some transitional time to explore the specific implementation of the law as large ambiguity was left when the law was passed. We use this temporal variable as a proxy for the motivation or eagerness of INGOs to register an RO, for some INGOs are inclined to avoid the cumbersome registration procedures at the early stage if there is no real consequences for continuing operations without registration, while others took the time to talk with multiple potential PSUs to find a more suitable superior or regulatory dynamic. (TODO: find a citation for this, probably Spires, if there is one.)

@@

There are various drawbacks to having to constantly file temporary activities. First would be high administrative costs. Since the filing of temporary activities needs to be submitted by INGOs’ Chinese partners, it would mean discussing and securing the partnerships 67 times, mostly with different organizations. Moreover, this model also imposes high uncertainty to the operation of INGOs. There is no guarantee whether the next Chinese partner would agree to collaborate and file the documentation for them. And it is impossible to run longer-term projects, given all temporary activities cannot last longer than a year. Finally, temporary activities would be programs Chinese partners feel comfortable to collaborate in, i.e., safe activities with little political and administrative risks. So there is less flexibility in the kinds of work an INGO can with temporary activities.

What factors could have contributed to the contrasting experiences of TNC and Greenpeace under the ONGO Law? After analyzing the data, the working approach and relationship with the government emerged as pivotal differences between the two INGOs, which fits existing discussions on the differential treatment of INGOs based on their relationship with the authority [TODO: add citations]. In a public interview with TNC’s RO in China about their registration experience, the chief representative stated TNC’s working approach in China has been “science-based”, using “non-confrontational, collaborative approach in working closely with governments, the business community and others” [@ChinaPhilanthropyReserchInstitute:2018a]. Over the years, they have provided technical support for the central government and local governments like the Yunnan and Zhejiang Province to build national parks and nature reserve areas and helped drafted such documents as the China Biodiversity Conservation Strategy and Action Plan (2011-2030). Using the non-confrontational working model and establishing long partnerships with the government enabled TNC to secure the consent of a national governmental agency in corresponding issue area to be their PSU, and thus successfully got registered and authorized to operate in a relatively large number of provinces. In contrast, Greenpeace states in their mission that they work as “independent campaigning organizations that use peaceful protest and creative confrontation to expose global environmental problems”. Their description of the work in China also uses such key words as “brings change”, “non-violent direct action”, and “impartiality and independence”. One project Greenpeace shares on their website was the exposure of the damages the company Sinar Mas Group's did in southwest Yunnan back in 2004. They conducted three rounds of  independent investigations and later revealed that the multinational paper industry company used reforestation as a cover while actually logging Yunnan's natural forests. Although this confrontation was not with the government, it showcases Greenpeace’s investigative, watchdog like work stance. To the authority, it would mean less predictable and manageable.

@@

@tbl-case-study summarizes the comparison of the cases of TNC and Greenpeace. The case study shows that in addition to the correlation of issue area and operational space found in the quantitative model, the dimension of relationship with the government also exerts influence on the contentious level and restrictions INGOs can face, revealing the potential patterns of differential treatment INGOs can receive in the *de facto* implementation of the law. Greenpeace also showed high organizational flexibility in the face of high legal barriers with the ONGO law : despite the various drawbacks discussed above, they were still able to keep conducting programs in China. A more evasive group of INGOs after the ONGO law would be those not having the same level of flexibility as Greenpeace and have left China either not choosing to seek registration or failed to do so and not able file temporary activities constantly. The variance in organizational flexibility can be worthy of further research.

@@

@@

>>>>

---
title: "Operational constraints and closing civic space: International NGO restrictions and China’s Overseas NGO law"
short-title: "China's ONGO Law and International Civil Society"
date: "2023-12-03"
author:
- name: Meng Ye
  affiliation: Georgia State University
  email: mye2@gsu.edu
  orcid: 0000-0002-6552-8338
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  TODO: **Update this based on the finished version of the paper.** China’s 2017 Overseas NGO (ONGO) Law is part of a larger global trend of legal restrictions on international nongovernmental organizations (INGOs). However, the effects of this global crackdown on INGOs have been difficult to measure since there is often divergence between the formal de jure regulations and the de facto implementation of these laws. Additionally, the severity and consistency of the enforcement of these laws often depends on the type of services provided by these organizations. In this paper, we explore the effect of China’s ONGO law on INGO operations in the five years since the passage of the law. We test whether and how INGO operational space under the ONGO law is influenced by organizations' issue areas. We argue that the less contentious the issue areas INGOs work in the more likely they are granted greater operational space, measured by the number of provinces they are legally authorized to work in. China provides an excellent setting for examining the phenomenon of closing space civic space, since the 2017 ONGO law hypothetically regulates both claim-making and service-providing NGOs, while the level of ambiguity of the law gives ample room for arbitrary discretion and conflicting attitudes by authorities. We test our argument by analyzing administrative data from 633 registered representative offices of INGOs registered in China between 2017–2021 and test our hypothesis with a multilevel Bayesian model. Our results speak to the broader literature on closing civic space and provide an empirical illustration of the practical effect of NGO restrictions on global civil society.
keywords:
  - international NGOs
  - civil society
  - authoritarianism
  - Chinese ONGO law
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/compassionate-clam>"
bibliography: references.bib
link-citations: true
---

@@

In the past two decades, nearly 100 countries have proposed or enacted legal measures designed to restrict nongovernmental organizations (NGOs) and civil society more broadly [@Chaudhry:2022; @ChaudhryHeiss:2021; @CarothersBrechenmacher:2014; @DupuyPrakash:2018; @DupuyPrakash:2020; @DupuyRonPrakash:2016].[^ngo-clarification] According to the CIVICUS Civil Society Monitor, as of 2023 only `r civicus_prop$true$prop_nice` of countries have open and unrestricted civic space [@ChaudhryHeiss:2022a; @CIVICUSMonitor]. This global trend of closing or shrinking civic space has affected countries of all regime types, including both autocracies and democracies such as Morocco and Palestine [@AtiaHerrold:2018], India [@press-trust-india:2019], Vietnam [@Sidel:2023], Canada [@CIVICUS:2013], and the United States [@Cramer:2020]. On April 28, 2016, China passed its own set of civil society restrictions with its Law of the People's Republic of China on Administration of Activities of Overseas Nongovernmental Organizations in the Mainland of China---or "2016 ONGO Law"---which formally took effect on January 1, 2017. The law requires all international NGOs (INGOs) to either establish formal in-country representative offices (ROs) or file short-term temporary activities to lawfully operate in mainland China. Without recognized legal status, INGOs are prohibited from working in the country, and even with legal recognition, INGO programming is limited to government-approved activities.

@@

By focusing on INGO operational space in China, we provide a unique empirical view into the *de facto* regulatory environment for foreign civil society in an authoritarian state. Many previous studies use broad indexes of civil society openness, such as the core civil society index from the Varieties of Democracy project [@LianMurdie:2023; @ChaudhryHeiss:2022]. These measures capture the general regulatory environment for all civil society---including domestic NGOs, INGOs, and informal organizations---but do not reflect the specific implementation of laws for INGOs. In this paper, though, we use a mixed methods approach to explore the regulatory environment for INGOs in particular, providing a much richer description and analysis of the impact of INGO laws.

Below, we provide a brief overview of China's 2016 ONGO Law and the different regulatory mechanisms---and practical limitations---it provides for INGOs to obtain legal access to the country. We then summarize existing research on authoritarian regulation of INGOs and the differential application of these laws based on regime preferences and organizational characteristics, and propose a set of empirical expectations based on this literature. We test these expectations with a mixed methods research design, combining Bayeisan analysis of administrative data from `r nrow(ongo)` formally registered INGOs with a comparative case study of Greenpeace and The Nature Conservancy, two INGOs working on similar issue areas which took different approaches to operating in China after 2017. We conclude with a broader discussion of authoritarian civil society regulations and propose future methods for exploring the practical operating environment for INGOs working in such regimes.

# The legal framework established by the 2016 ONGO law

International NGOs began to play an important role in China in the early 1990s as part of the country's Reform and Opening-Up policy, and organizations such as Save the Children and the Ford Foundation established local offices. Since then, INGOs and global civil society have played pivotal roles in providing funding, organizational and program management skills, as well as catalytic support for the emergence of China's domestic nonprofit sector [@Yin:2009]. Prior to the implementation of the 2017 ONGO Law, regulation of INGOs in China was ambiguous [@Shieh:2018; @Spires:2022]. The only legal document relevant to INGOs was a section of the Regulation on the Management of Foundations that specifies the registration of ROs of foreign foundations with the Ministry of Civil Affairs, the regulator of Chinese domestic nonprofits. However, the actual application of the regulation was very limited: until the ONGO Law took effect in 2017, there were fewer than 30 foreign foundation ROs registered, while some Chinese officials estimated in 2016 that there were roughly over 7,000 INGOs operating in China [@YeHuang:2018; @Spires:2020].

Many INGOs gained legal status by registering as foreign-invested companies [@Sidel:2016], for the administrative convenience such as hiring employees [@HolbigLang:2022; @Yin:2009]. However, such a legal form is not quite compatible with the non-profit-distributing nature of INGOs. Chinese companies are intended to be for-profit only, while nonprofits are regulated under other legal forms, such as foundations and social service organizations [@Ye:2021]. Thus, INGOs registered ad foreign-invested companies had to navigate through for-profit institutional logics, with doctrines like protecting shareholders' investment returns.

The legal framework governing INGOs was overhauled by the 2016 ONGO Law. The new law applies to all overseas (including Hong Kong, Macau, and Taiwan) NGOs that conduct not-for-profit activities in the fields of economic development, education, science, technology, culture, public health, sports, environment protection, poverty alleviation, and disaster relief. The ONGO Law provides two pathways for INGOs to lawfully operate in China (see @fig-ro-ta-pathways): (1) submitting an application for formal permanent registration as ROs, or (2) filing documentation of temporary activities with the Public Security Bureaus. Formal registration as an RO allows INGOs to maintain an ongoing presence in the country, while temporary activities are intended for INGOs that need to conduct one-off programs that last no longer than one year. Using one of these two channels to obtain legal status became mandatory in 2017.

@@

ro_nodes <- tibble(
  step = c(
    "Find a professional supervisory unit (PSU)",
    "Obtain PSU approval",
    "Open organization bank accounts",
    "Submit annual work plans to PSU and Public Security Bureau",
    "Keep programming within approved scope",
    "Submit annual reports to PSU and Public Security Bureau"
  ),
  type = "Representative offices"
)
ro_edges <- tibble(from = 1:5, to = 2:6)
ro_graph <- tbl_graph(nodes = ro_nodes, edges = ro_edges)

ta_nodes <- tibble(
  step = c(
    "Find an eligible Chinese partner to file paperwork",
    "File activity information with Public Security Bureau",
    "Keep programming within approved scope",
    "Finish work within a year",
    "Submit final project report to Public Security Bureau"
  ),
  type = "Temporary activities"
)
ta_edges <- tibble(from = 1:4, to = 2:5)
ta_graph <- tbl_graph(nodes = ta_nodes, edges = ta_edges)

@@

Registering as an RO under the ONGO Law is a complex process that requires the consent of a Chinese Professional Supervisory Unit (PSU)---a governmental or quasi-governmental unit related to an INGO's issue area which oversees and co-supervises the INGO with Public Security Bureaus. Obtaining the consent of a PSU to apply for registration has become a key challenge for INGOs interested in formal registration, since potential PSUs do not have a legal obligation to assume such a demanding role [@Jia:2017; @Shieh:2018; @YeHuang:2018]. Organizations vary in their ability to attract partner PSUs, and INGOs that are unable to find suitable PSUs can end up not being able to registered or registered with great limitation in terms of work areas and geographical reach. INGOs are authorized to operate within China on a province-by-province basis. Many organizations are allowed to work in all 32 mainland provincial administrative regions, but many are also only authorized to work in a handful, or even only a single province. Similarly, INGOs need to find an eligible Chinese partner, including governmental, quasi-governmental and non-profit organizations to file temporary activities for them. After successfully gaining legal status, INGOs also need to abide by the operational regulations by the law, such as no fundraising within mainland China, only channeling the funds through RO's registered account or the Chinese partners' accounts, filing reports regularly and stay within the geographical and issue scopes as approved.

@@

The authoritarian benefits of civil society regulations go beyond simply allowing or disallowing specific INGOs. The flexible *de facto* implementation of formal civil society laws allows authoritarians to "use nonprofit regulation as a tool of political control to shore up their continued rule" [@Spires:2020, 573], providing them with an additional policy tool for their stability-seeking calculus. These laws allow governments to capture and reshape INGO programming in their favor. @LianMurdie:2023 illustrate how increased civil society restrictions have counter-intuitively led to an increase in cooperative, regime-friendly INGO programming, with INGOs praising government policies, engaging in joint programming with the government, and changing longstanding programs to align with government desires. China's ONGO law illustrates this dynamic well. The ONGO law requires that INGOs partner with government agencies and find collaborating PSUs, which provides the government with more influence over INGO decision making. For example, Oxfam gained RO status in 2017, but only after negotiations with its proposed PSU forced it to stop collaborating with other organizations and local agencies [@Li:2020]. In 2018, @BatkeHang:2018 posited that the ONGO law was showing signs of reorienting the missions of legally recognized INGOs and the overall scope of international civil society to be more explicitly aligned with government preferences. Holbig & Lang confirmed this prediction, finding that the issue areas addressed by legally recognized INGOs had "shift\[ed\] towards fields of activity high up the government's domestic policy agenda" [-@HolbigLang:2022, p. 587]. Civil society laws thus filter out organizations opposed to the host regime's policy preferences and grant increased government control over safer and more beneficial INGOs.

@@

To test the relationships between government preferences, INGO issue area, and INGO operational space, we apply this framework to China's ONGO law. Previous research has looked at the practical implementation of civil society laws on a global scale [@LianMurdie:2023; @Chaudhry:2022], which naturally blurs the specific policy mechanisms regimes use for controlling and leveraging global civil society. By analyzing the population of formally registered INGOs between 2017–2021, we can more precisely explore how a single formal anti-INGO law has shaped the operational space for foreign organizations working in the country. Specifically, we can observe exactly how the Chinese government has used the RO registration process to strategically limit the potential reach of INGOs and reshape their issue areas. We use the number of provinces that INGOs are authorized to work in as a proxy for INGO operational space based on the assumption that INGOs need to or hope to be able to carry out their programs in as broad a geographical scope as possible [@Li:2020]. For example, in emergency circumstances like the COVID-19 pandemic, it would be extremely difficult for INGOs to plan beforehand which areas might need their service. Having the potential to shift programs to other provinces as needed thus provides a form of operational space.

@@

We examine these theoretical expectations using a mixed methods approach. First, we explore a rich dataset of all ROs registered in China between 2017–2021 using a combination of descriptive statistics and Bayesian models to understand how the international civil society in China has been reshaped since the implementation of the ONGO law. We then complement this statistical analysis with a comparative case study of two INGOs---The Nature Conservancy and Greenpeace ---to explore in detail how the ONGO has specifically influenced organizational operational space.

@@

Our data comes from INGOs' public registration documents disclosed by the Ministry of Public Security's Overseas NGO Office, which maintains a website with complete documentation of all the ROs and temporary activities registered or filled across China. We include all INGO registrations between January 1, 2017, when the law took effect, to December 31, 2021, with information for all `r nrow(ongo)` actively registered organizations. The filing documents for ROs disclose each INGO's name, address, registration date, regulating Public Security Bureau, PSU, issue area and purposes of the RO, and the specific provinces they can work in. We merge this INGO data with translated English organization names data from ChinaFile. Our data and reproducible code is available at `ANONYMIZED_DOI_URL`.

@@

Our dependent variable is the count of provinces that INGOs are authorized to work in. We derive this value from the official registration record, which lists all the provincial regions each INGO RO can conduct activities in. This value ranges from 1–32 (the total number of province-level jurisdictions in mainland China).

Our key independent variables correspond to each of our three theoretical expectations. First we code each INGO's issue area based on its mission statement reported in its registration documentation. Given there is certain level of subjectivity in the coding process, we cross-checked our manual coding of issue areas---Author 1 assigned each organization into one of the issue areas defined in Article 3 of the ONGO Law, and Author 2 cross-checked the coding for ambiguous cases. We include our coding rules for overlapping issue areas in the appendix. We categorize each organization into nine different issue areas (see @tbl-issues).

@@

Our second independent variable is a dichotomous indicator of whether the INGO explicitly states that it intends to operate locally or sub-provincially. There are cases where the aim of the INGO is attached to a locality, either having a county or city name in their RO name or purposes, or the INGO is actually set up by people in China who registered the RO to communicate between the "hometown" and the INGO set up abroad.

Our final independent variable measures how much time elapsed between January 1, 2017 (when the law took effect) and the organization's official registration date. There is great variance in the timing of registering an RO among INGOs, since there has been a *de facto* grace period for INGOs to seek legal presence after the law took effect [@Shieh:2018]. This is largely because both INGOs and the regulators needed some transitional time to explore the specific application of the law as large ambiguity was left when the law was passed. We use this temporal variable as a proxy for the motivation or eagerness of INGOs to register an RO, for some INGOs are inclined to avoid the cumbersome registration procedures at the early stage if there is no real consequences for continuing operations without registration, while others took the time to talk with multiple potential PSUs to find a more suitable superior or regulatory dynamic. [@Jia:2017a]

@@

There are various drawbacks to having to constantly file temporary activities. First would be high administrative costs. Since the filing of temporary activities needs to be submitted by INGOs’ Chinese partners, it would mean discussing and securing the partnerships 67 times, mostly with different organizations. Moreover, this model also imposes high uncertainty to the operation of INGOs. There is no guarantee whether the next Chinese partner would agree to collaborate and file the documentation for them. And it is impossible to run longer-term projects, given all temporary activities cannot last longer than a year. Finally, temporary activities would be programs Chinese partners feel comfortable to collaborate in, i.e., safe activities with little political and administrative risks. So there is less flexibility in the kinds of work an INGO can do with temporary activities.

What factors could have contributed to the contrasting experiences of TNC and Greenpeace under the ONGO Law? After analyzing the data, the working approach and relationship with the government emerged as pivotal differences between the two INGOs, which fits existing discussions on the differential treatment of INGOs based on their relationship with the authority [@Plantan:2022]. In a public interview with TNC’s RO in China about their registration experience, the chief representative stated TNC’s working approach in China has been “science-based”, using “non-confrontational, collaborative approach in working closely with governments, the business community and others” [@ChinaPhilanthropyReserchInstitute:2018a]. Over the years, they have provided technical support for the central government and local governments like the Yunnan and Zhejiang Provinces to build national parks and nature reserve areas and helped drafted such documents as the China Biodiversity Conservation Strategy and Action Plan (2011-2030). Using the non-confrontational working model and establishing long partnerships with the government enabled TNC to secure the consent of a national governmental agency in corresponding issue area to be their PSU, and thus successfully got registered and authorized to operate in a larger number of provinces. In contrast, Greenpeace states in their mission that they work as “independent campaigning organizations that use peaceful protest and creative confrontation to expose global environmental problems”. Their description of the work in China also uses such key words as “brings change”, “non-violent direct action”, and “impartiality and independence”. One project Greenpeace shares on their website was the exposure of the damages the company Sinar Mas Group's did in southwest Yunnan back in 2004. They conducted three rounds of  independent investigations and later revealed that the multinational paper industry company used reforestation as a cover while actually logging Yunnan's natural forests. Although this confrontation was not with the government, it showcases Greenpeace’s investigative, watchdog like work stance. To the authority, it would mean less predictable and manageable.

@@

@tbl-case-study summarizes the comparison of the cases of TNC and Greenpeace. The case study shows that in addition to the correlation of issue area and operational space found in the quantitative model, the dimension of relationship with the government also exerts influence on the contentious level and restrictions INGOs can face, revealing the potential patterns of differential treatment INGOs can receive in the *de facto* implementation of the law.

@@

Greenpeace also showed high organizational flexibility in the face of high legal barriers with the ONGO law : despite the various drawbacks discussed above, they were still able to keep conducting programs in China. A more evasive group of INGOs after the ONGO law would be those not having the same level of flexibility as Greenpeace and have left China either not choosing to seek registration or failed to do so and not able file temporary activities constantly. The variance in organizational flexibility can be worthy of further research.

@@

## Coding rules for issue ares

1. Education vs. Charity
If charitable activities are all education related, e.g. building schools, donating education facilities and books, coded as education
If the aids provided include other community support, e.g. building bridges, roads, medical facilities, coded as charity

2. Economy and trade vs. Industry association
If the trade promoted is focused on a particular industry (poultry, grains etc.), coded as industry association

3. Arts and culture 
Include sports organizations too.

4. Industry association vs. health, science and technology, arts and culture

There are professional association on health providers and certain techniques. Coded as health, science & technology if work areas go beyond serving members only: for example conducting activities of promoting and training about certain techniques.

If specialized association only serves the member companies: coded as industry associations

5. Health vs. Charity 
If only providing medical aids to underprivileged communities, not building schools etc., coded as health

6. Charity vs. General 
If multiple work, e.g., poverty alleviation, education, health can be encompassed by Charity and humanitarian, then the former

7. Multiple ROs are coded based on the overall work area spans of the INGO, not the specific RO, so to ensure the issue area coding is consistent across ROs
